Comparing variations in sediment transport capacity between subglacial and subaerial systems

### 2023-03-14 theory.tex
Changes to key points and abstract...

\paragraph{Key findings:}
\begin{enumerate}
\item The slowly evolving size of subglacial channels means that variations in water discharge must be accommodated by changing the velocity of water flowing through the channel, as opposed to increasing channel area.

\item Increased variability in water velocity results in greater variability in sediment transport capacity in subglacial systems compared to subaerial ones, although the variable width of subaerial channels reduces some of the total sediment transport capacity across the channel.

\item  As water discharge changes, differences in the sediment transport capacity response of subaerial and subglacial channels could impact sediment transport processes at proglacial margins as well as the interpretation of sediment transport records in glacial environments. 
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which the sediment is mobilized.
In subaerial channels, changing water discharge can be accommodated by changes in the water depth, the width of the channel that  water flows over, and water velocity.
In subglacial channels, however, water flows through a conduit with a slowly evolving size, because the water is pressurized by the glacier ice.
As a result, changing water discharge may modify the water velocity, but not the  shape of the channel.
Parameterizations of channel dynamics indicate that greater variability in sediment transport capacity in subglacial systems occurs across a wide range of channel geometries and flow conditions.
The increased variability in water velocity and sediment transport capacity impacts sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.

@@

\laysummary % 200 words
In both subglacial and subaerial channels, sediment transport capacity depends, in the first order, on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom, depending largely on the water velocity.
In subaerial channels, changing water discharge will change the water depth, the water velocity (controlling sediment transport capacity),  and channel width (which dictates the span over which sediment transport  will occur).
Conversely, in subglacial environments, on short times scales of hours to days, subglacial channels only experience a change in water velocity following changing water discharge, because the water flows through a subglacial conduit with a fixed geometry.
The fixed conduit geometry also means that the channel width, over-which sediment can be mobilized, does not respond to changing water discharge.
Parameterizations of the processes show that sediment transport capacity varies far more in subglacial channels compared to subaerial ones, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
This increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.

@@

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021} and the Andes \citep{vergara2022}.
  In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \citep{bendixen2017,costa2017,li2021}.
  In other environments, enhanced access of melt water to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
  The observed changes to these systems create an imperative to examine the processes controlling sediment discharge from glacierized catchments. 
  Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not contrasted with the subaerial river systems that these glaciers feed.

  Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}. 
  On shorter time periods, pressurized water transports sediment from the glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e. in a transport-limited regime).

@@

  Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
  In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
  The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
  As a result, the subglacial channel size evolves over weeks or months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.

  As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these environments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacier compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by thick black line. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
        Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methods}
  \label{sect:meth}
    Parameterizations below represent relationships among water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Fig. \ref{fig:cartoon}).
  The evaluation of shear stress  omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.

@@

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where $v_g = \frac{Q_w}{S_g}$ is the water velocity.
  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \citep{rothlisberger1972,delaney2019}.

  For purposes discussed below, width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

@@

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v$, $v_f$), shear stress ($\tau$, $\tau_f$), and integrated shear stress ($\tau_t$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.

  \section{Results}
  First, the subglacial hydraulics parameterization is applied to a representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$.
  Both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge. The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6-- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.

@@

  The direct comparison between the systems remains difficult as they have with different hydraulic potentials, and the response of channel width to water discharge varies considerably based on the values chosen for channel shape factors $\beta$ and $k$.

  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
  While the subglacial channel geometry need not be in a steady state for the water to remain pressurized, the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.

  In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might slightly modify the width of the subglacial channel.

@@

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems or partially supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment frequently  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.

@@

  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
  Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier. The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
  Note that the transition from pressurized to open channel flow could occur above glacier termini when cavity closure rates are reduced and water can flow freely \citep{egli2021b}. 
  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow near the glacier terminus \citep{perolo2018}.

  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie \citep[c.f.][]{swift2005,riihimaki2005,costa2017,vanas2017,hasholt2018}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.

  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport parameters (Figure~\ref{fig:range}\, d,\,e,\,f).
  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,hasholt2018}.
  Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in these catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:range}~ and~\ref{fig:model_outs}).

@@

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
  This sediment mobilization and deposition would occur in response to both evolving channel geometry and water discharge with accumulating with melt down-glacier \citep{beaud2018,delaney2019}.
  Increasing variability in sediment transport down-glacier is likely because of spatial discontinuities in sediment availability, increasing water discharge down the glacier, and the transport times of subglacial sediment \citep{williams1989,delaney2019}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system, for instance, compounds the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, variations could be the result of increased water velocity resulting from reduced channel size, not water discharge, therefore increasing sediment transport capacity.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
  For instance, sediment expelled from a glacier in a transport-limited regime reacts, in addition to water discharge, to sediment discharge capacity defined by the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}.
  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over long time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}. 
  Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion sliding and water pressure variations  \citep{iverson2012,herman2015}.

  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, in addition to the stochastic nature of erosion and deposition in fluvial evironments \citep{castletort2003,jerolmack2010,romans2016}.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly in part because sediment transport capacity relies on the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
  This suggests that a larger climate perturbation might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.

@@

  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond change water discharge.
  The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
  Even so, variability in sediment transport capacity across a channel's width is consistently higher in subglacial channels, compared to subaerial ones.

@@

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}

>>>>

\paragraph{Key findings:}
\begin{enumerate}
\item Variations in water discharge in subglacial systems are accommodated by water velocity of water, given the slow evolution of channel size.

\item  Greater variability in water velocity in subglacial systems causes sediment transport capacity to vary more compared to subaerial systems.

% although the evolving width of subaerial channels tempers some difference in total sediment transport capacity between the two systems.

\item The systems' divergence in transport capacity response impacts sediment dynamics at glacier margins and interpretation of sediment records.
\end{enumerate}

\abstract % 150 words
Sediment transport capacity depends mainly on the shear stress or velocity of water flowing through a channel and the channel width over which to mobilize sediment.
In subaerial channels, changing water discharge can be accommodated by rapid changes in the width of the channel over which water flows, the water depth, and the water velocity.
In subglacial channels, however, water flows through a conduit whose size evolves  slowly, because the water is pressurized by the glacier ice above.
As a result, variations in water discharge may modify the water velocity, but not the channel's width.
Here, parameterizations of subglacial and subaerial channels indicate that sediment transport capacity varies more in subglacial systems compared with subaerial systems, across a wide range of channel geometries and flow conditions.
Large variations in subglacial water sediment transport capacity may impact sediment transport processes and the interpretation of records of sediment transport from glacierized catchments.

@@

\laysummary % 200 words
In both subglacial and subaerial channels, sediment transport capacity depends mainly on 1) the channel width over which the sediment is mobilized  and 2) the shear stress applied by flowing water on sediment at the channel bottom; the shear stress depends largely on the water velocity.
In subaerial channels, changing water discharge will  change the water depth,  the channel width (which dictates the distance  over which sediment transport  occurs), and the  water velocity (which controls the sediment transport capacity).
 In subglacial environments, however, on times scales of hours to weeks, changing water discharge will only alter  water velocity, not channel area, because the subglacial conduit maintains a largely fixed geometry.
The fixed conduit geometry also means that the channel width over which the sediment is  mobilized also does not respond to changing water discharge.
Parameterizations of these processes show that sediment transport capacity varies far more in subglacial systems compared to subaerial systems, even though changing channel width can accommodate some sediment transport variability in subaerial channels.
The increased variability in sediment transport capacity in subglacial systems, compared to subaerial ones, has several important implications for sediment transport in glacierized catchments and the interpretation of sediment discharge records.

@@

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Increases in sediment transport  have been observed in Greenland \citep{bendixen2017}, the European Alps \citep{costa2017}, the Himalayas \citep{li2021}, and the Andes \citep{vergara2022}.
  In some of these regions, increased water discharge and glacier melt have been attributed to the greater sediment transport \citep{bendixen2017,costa2017,li2021}.
  In other cases, enhanced access of meltwater to subglacial sediment at high elevations was attributed as the cause of increased sediment transport \citep{delaney2020,vergara2022}.
  Observed changes to sediment transport in glacierized catchments create an imperative to examine the processes controlling sediment discharge here. 
  Yet, the relationship between water discharge and sediment transport in subglacial environments remains minimally discussed and not contrasted with the subaerial river systems that these glaciers feed.

  Over long periods, processes such as glacier abrasion and quarrying sculpt landscapes and create sediment to be transported fluvially from below glaciers \citep[c.f.][]{hallet1979,iverson2012,ugelvig2018}. 
  On shorter time periods, pressurized water transports sediment from the below glacier \citep{walder1994,creyts2013,beaud2018}, should enough sediment be present below the glacier (i.e. in a transport-limited regime).

@@

  Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
  In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
  The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
  As a result, the subglacial channel size evolves over days to months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.

  As a result, sediment mobilization in subaerial and subglacial systems responds differently to changing water discharge.
  These differences have been implicitly included in a wide range of available models that  quantify sediment transport from these catchments in both subglacial and subaerial systems  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
  The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016}.
  Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial compared to subaerial systems \citep[e.g.][]{muller1968,richards2003,ganti2016}.
  Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial systems has been minimally discussed.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, systems are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$, denoted by the thick black line. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
        Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

  \section{Methods}
  \label{sect:meth}
    Parameterizations below represent relationships amongst water discharge, subglacial water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
  Because sediment transport largely depends on shear stress applied by water flow \citep{shields1936}, both parameterizations calculate this value and integrate  it across the channel bed (Figure \ref{fig:cartoon}).
  The evaluation of shear stress omits a grain size parameter and, thus, leaves open the selection of a sediment transport relationship  \citep[e.g.][]{shields1936,meyer1948} to  calculate sediment transport capacity.
  As a result, relative relationships in shear stress, as opposed to sediment transport capacity, are examined.

@@

  The shear stress, $\tau$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
  \begin{equation}
    \label{eq:tau}
    \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
  \end{equation}
  % 
  where  the water velocity $v_g$ is $v_g = \frac{Q_w}{S_g}$.
  Here, $Q_w$ represents the instantaneous water discharge, rather than $Q_w^*$, which is used to evaluate the channel geometry.
  With this formulation, the processes in an R-channel, through which water flows subglacially, can be represented algebraically \citep{rothlisberger1972,delaney2019}.

    Width integrated shear stress is represented as $\tau_{tg}=w_{cg}\,\tau_g $.

@@

  The variance $ \chi$ of a variable $\Gamma$ is given by 
  \begin{equation}
    \label{eq:var}
    \chi(\Gamma) \,=\, 1 - \frac{\mathrm{min}(\Gamma)}{\mathrm{max}(\Gamma)}\,\,.
  \end{equation}
  % 
  \noindent Below we examine variables, $\Gamma$, as water discharge ($Q_w$) water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and integrated shear stress ($\tau_{tg}$, $\tau_{tf}$) with the purpose of evaluating total shear stress across the glacier bed.

  \section{Results}
  Initially, both parameterizations are forced with a water discharge ($Q_w$) varying between $8$ and $16$ \,\unit{m}$^{3}$\,\unit{s}$^{-1}$, to represent a range of sediment transport conditions, for instance over diurnal fluctuation in water discharge. The subaerial channel width factor $k$ is $3$\,\unit{m}$^{2}$\,\unit{s}$^{-1}$, and the Hooke angle $\beta$ is $\frac{\pi}{2}$.
  Values of Hooke angle $\beta$ and subglacial friction factor $f_r$ are chosen so that they produce reasonable values of water velocity ($0.6- 1.2$\,\unit{m}\,\unit{s}$^{-1}$) in accordance with measurements made through dye-tracing experiments to examine subglacial hydraulic conditions \citep[Section~\ref{sect:sub_mode}, Figure~\ref{fig:model_outs}; e.g.][]{werder2010}.
  The subglacial representative water discharge ($Q_w^*$; Equation~\ref{eq:DW}) of $12$\,\unit{m}$^{3}$\,\unit{s}$^{-1}$ is applied to establish the hydraulic diameter and shape of the channel.

@@

  The subglacial parameterization assumes that subglacial water is pressurized, which is a commonly held assumption in subglacial hydrology, though observations suggest that this may not always be the case \citep[e.g.][]{gimbert2016}.
  The subglacial channel geometry need not be in equilibrium with the hydrological conditions for the water to remain pressurized.
  Yet the subglacial parameterization assumes that the channel geometry must vary minimally over the timescales of varying  water discharge \citep[e.g.][]{nanni2020}.

  In R-channel formulations, peaks in the hydrograph could cause additional opening of the conduit due to pressure melt, thus causing a decrease in velocity compared to the results here \citep{rothlisberger1972}.
  A similar effect could occur with creep closure of the subglacial channel at the hydrograph minima.
  Additionally, at high water pressures, glacier uplift and bed separation \citep{andrews2014} might modify the width of the subglacial channel.

@@

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.

@@

  Increased variability in sediment transport may also impact sediment export from the subglacial environment into proglacial areas \citep[e.g.][]{delaney2017,perolo2018}.
  Here, velocities at the highest water discharges could generally decrease (Figure~\ref{fig:model_outs}, blue and green horizontal bars), in the transition of pressurized subglacial flow to subaerial open-channel flow as the water leaves the glacier. The decrease in water velocity potentially results in sediment deposition once sediment enters the proglacial area.
  Periodic sediment deposition in sediment deposition may even result in the transition between pressurized and subaerial flow below the glacier terminus \citep{perolo2018}, especially when cavity closure rates of the ice are slow \citep{egli2021b}.

  In large catchments and further downstream of proglacial margins, water discharge is generally less variable than at the tops of catchments, where glaciers lie in alpine settings \citep[c.f.][]{swift2005,riihimaki2005,costa2017}.
  Here, external hydrological factors further downstream of glaciers could result in smaller variations in sediment transport capacity in many subaerial systems compared to subglacial ones.

  The parameterization here shows that the smallest variability in water discharge has the largest difference in relative variability in sediment transport parameters (Figure~\ref{fig:range}\, d,\,e,\,f).
  At locations such as the Greenland Ice Sheet, smaller diurnal variations in water discharge compared to many alpine catchments \citep[c.f.][]{swift2005,riihimaki2005,vanas2017,hasholt2018}.
  Therefore, the variability in sediment transport capacity in the transition from the ice sheet to the proglacial river could be more dramatic in Greenland than in alpine systems as water leaves the glacier and moves downstream.
  Thus the disparity in sediment transport capacity between the subglacial and subaerial systems could be most pronounced in larger catchments, potentially resulting in temporally mismatched periods between when glaciers introduce sediment to these river systems  and when the river mobilizes it (Figures~\ref{fig:model_outs}~ and~\ref{fig:range}).

@@

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system may compound the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.

  Additional processes and erosional hiatuses may further complicate signals of sediment discharge from glaciers in response to climate, compared to subaerial systems \citep{jansson2005,ganti2016}. 
  For instance, sediment expelled from a glacier in a transport-limited regime reacts, in addition to water discharge, to sediment discharge capacity defined by the thickness of the ice, controlling the closure of the channel, and the surface slope of the glacier, controlling water velocity \citep[Section~\ref{sect:sub_mode}; ] []{rothlisberger1972,shreve1972,delaney2022,stevens2022}.
  Conversely, sediment transport in most subaerial systems increases with water discharge and bed slope, with the bed slope, probably being more stable over yearly to century-scale time periods \citep[Section~\ref{sect:fluv}; e.g.][]{muller1968,whipple1999,wong2006,wickert2019}.

  Records of sediment discharge have been used to establish the relationship, or lack thereof, between sediment discharge and climate in glacial systems \citep[e.g.][]{koppes2009a,willenbring2016,mariotti2021}.
  The variability in subglacial sediment transport capacity presented here applies most generally to short-time scales responsible for the size of subglacial channels.
  Yet, identifying climatic signals in sediment transport from transport-limited glaciers may require higher thresholds of climatic perturbation compared to subaerial systems \citep{tofelde2021}, in addition to the stochastic nature of erosion and deposition in fluvial environments \citep{castletort2003,jerolmack2010,romans2016}.
  For instance, sediment discharge records from two glaciers in the Swiss Alps show that $40$ to $50$\% of a season's sediment discharge occurs when water discharge is below the $75^{\mathrm{th}}$ quantile of the season \citep{delaney2018}, possibly in part because sediment transport capacity responds the size of the glacier conduit, in addition to water discharge (Section~\ref{sect:sub_mode}).
  This suggests that a larger  perturbation of water discharge might be needed to significantly alter the sediment discharge capacity from the glaciers compared to subaerial systems, to overcome the noise resulting from increased variability in sediment transport capacity.
Furthermore, if no sediment is available and the glacier is in a supply-limited case, then the sediment discharge record will also represent additional processes of bedrock erosion from sliding and water pressure variations  \citep{iverson2012,herman2015}.

@@

  Parameterizations of subglacial and subaerial water flow show that subglacial systems exhibit increased variability in sediment transport capacity as they respond  to changes in water discharge.
  The variability in sediment transport capacity is reduced when the evolving channel width is accounted for in subaerial systems.
  Even so, variability in sediment transport capacity across a channel's width is consistently higher in subglacial channels, compared to subaerial ones.

@@

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and subaerial systems. Across the parameter space in Figure~\ref{fig:range}. }
    \label{fig:gammas}
  \end{figure}
\end{center}
